---
title: "Hepatitis C & Hepatitis B in Saudi Arabia"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In 2019, we estimated the relative risk of hepatitis C and hepatitis B across various regions of Saudi Arabia. For hepatitis C, the highest relative risks are observed in the western regions, particularly in Makkah and Al Bahah, while the northern regions, such as Al-Jawf and Tabuk, exhibit the lowest risks. In contrast, for hepatitis B, the southern regions, particularly Jazan, display the highest relative risk, with values indicated by the yellow color corresponding to approximately 2.5. Meanwhile, the northern regions, such as Al-Jawf and Hail, show the lowest relative risks for hepatitis B.


```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA}




# The map below illustrates the standardized incidence ratios (SIR) for hepatitis C and hepatitis B in Saudi Arabia in 2019. The spatial patterns for many regions appear to be highly similar, which may be attributed to the shared risk factors for both diseases

library(sf)
library(SpatialEpi)
library(sp)
library(raster)
library(readxl)
library(mapview)
library(dplyr)
library(ggplot2)
library(patchwork)
library(viridis)
library(openair)
library(terra)
library(leaflet)
library(tigris)
library(pander)
library(spdep)
library(INLA)
library(gridExtra)
#library(rgeoboundaries)
library(rnaturalearth)
library(htmltools)

#######################
# map of Saudi Arabia
#######################


map <- ne_states(country = "Saudi Arabia", returnclass = "sf")
map <- st_set_crs(map, 4326) 
map <- map[order(map$name),]
maplatlong <- st_transform(map, 4326)

# Change the names to match the data of hepatites names

map$name[1]="Asir"
map$name[2]="Al Bahah"
map$name[3]="Northern Borders"
map$name[4]="Al Jawf"
map$name[5]="Madinah"
map$name[6]="Qassim"
map$name[7]="Riyadh"
map$name[8]="Eastern Province"
map$name[9]="Hail"
map$name[10]= "Jazan"
map$name[11]="Makkah"
map$name[12]="Najran"
map$name[13]="Tabuk"



setwd("~/Desktop/files/hepatite/dash/dash")

####################
# Hepatitis Data
####################
data <- read.csv("~/Desktop/files/hepatite/dash/dash/data/hepatites.csv", header = T)




###########################################
# Merge the data of hepatitis with the map
###########################################





map <- merge(map, data, by.x = "name", by.y = "regions")



###################################
# Plot the cases of Hepatitis Data
###################################








################### 
# Expected
##################


map$E.c <- expected(
  population = map$population2019 ,
  cases = map$HepatitisC, n.strata = 1
)



map$E.b <- expected(
  population = map$population2019 ,
  cases = map$HepatitisB, n.strata = 1
)



##################
# SIR
##################

map$SIR.c <- map$HepatitisC/map$E.c

map$SIR.b <- map$HepatitisB/map$E.b



# Set the color scale limits based on the min and max SIR values across both datasets
common_limits <- range(c(map$SIR.c, map$SIR.b), na.rm = TRUE)
boundaryregion <- st_union(map)

# Create a common color palette for both datasets based on the common limits
palette <- colorNumeric(palette = "viridis", domain = common_limits, na.color = "transparent")

# Create the leaflet map for SIR.c
leaflet_sc <- leaflet(map) %>%
  addTiles() %>%
  addPolygons(data = map, 
              fillColor = ~palette(SIR.c), 
              fillOpacity = 0.7, 
              color = "gray", 
              weight = 1,
              label = ~paste0("SIR (C): ", SIR.c)) %>%
  addLegend(pal = palette, values = ~SIR.c, title = "SIR (C)")

# Create the leaflet map for SIR.b
leaflet_sb <- leaflet(map) %>%
  addTiles() %>%
  addPolygons(data = map, 
              fillColor = ~palette(SIR.b), 
              fillOpacity = 0.7, 
              color = "gray", 
              weight = 1,
              label = ~paste0("SIR (B): ", SIR.b)) %>%
  addLegend(pal = palette, values = ~SIR.b, title = "SIR (B)")

# Display the maps



# browsable(
#   tagList(
#     div(
#       style = "display: inline-block; width: 48%; margin-right: 3%;", 
#       leaflet_sc
#     ),
#     div(
#       style = "display: inline-block; width: 48%;", 
#       leaflet_sb
#     )
#   )
# )



###################################################
# Joint Model between Hepatitis C and Hepatitis B
###################################################

nb <- poly2nb(map)
nb2INLA("map.adj", nb)
g <- inla.read.graph(filename = "map.adj")



b = length(map$HepatitisC)
y = c( map$HepatitisC, map$HepatitisB)
int = c(rep(1 ,13),rep(2,13))
S= c(1:26)
b1 = c(1:13,rep(NA ,13))
b2= c(rep(NA,13),1:13)
m = as.factor(int)


E1 = c(map$E.c,rep(NA,b))
E2 = c(rep(NA,b) , map$E.b)



formula = y ~  -1  + m + offset(log(E1))+ offset(log(E2))+ 
  f(S, model="besag2", graph=g, scale.model=TRUE)+
  f(b1 , model = "bym", graph=g, scale.model=TRUE)+
  f(b2 , model = "bym", graph=g, scale.model=TRUE)

r <- inla(
  formula,
  family = "poisson",
  data = data.frame(y, S, m, b1, b2),
  verbose = TRUE,
  control.predictor = list(compute = TRUE),
  control.compute = list(return.marginals.predictor = TRUE)
)


# pander(round(r$summary.hyperpar[,c(1,2,3,5)],3))
# 
# pander(r$summary.fixed[,c(1,2,3,5)])


# The relative risk for Hepatitis C from the joint model


map$RRC <- r$summary.fitted.values[c(1:13), "mean"] / map$E.c
boundaryregion.c <- st_union(map)

# The relative risk for Hepatitis B from the joint model

map$RRB <- r$summary.fitted.values[c(14:26), "mean"] / map$E.b








# Calculate the common range for the color scales
common_range <- range(c(map$RRC, map$RRB))

# # Plot for RRC
# C <- ggplot(data = boundaryregion.c) + 
#   geom_sf() +
#   geom_sf(data = map, aes(fill = RRC)) +
#   scale_fill_viridis_c(option = "viridis", name = "RR", limits = common_range) +  # Use common range
#   theme(plot.title = element_text(hjust = 0.5, size = 16))  # Center and enlarge the title
# 
# # Plot for RRB
# B <- ggplot(data = boundaryregion.c) + 
#   geom_sf() +
#   geom_sf(data = map, aes(fill = RRB)) +
#   scale_fill_viridis_c(option = "viridis", name = "RR", limits = common_range) +  # Use common range
#   theme(plot.title = element_text(hjust = 0.5, size = 16))  # Center and enlarge the title
# 
# # Plot both
# C + B



# Create a common color palette for both datasets based on the common limits
palette <- colorNumeric(palette = "viridis", domain = common_range, na.color = "transparent")

# Create the leaflet map for SIR.c
leaflet_scr <- leaflet(map) %>%
  addTiles() %>%
  addPolygons(data = map, 
              fillColor = ~palette(RRC), 
              fillOpacity = 0.7, 
              color = "gray", 
              weight = 1,
              label = ~paste0("RR (C): ", RRC)) %>%
  addLegend(pal = palette, values = ~RRC, title = "RR (C)")

# Create the leaflet map for SIR.b
leaflet_sbr <- leaflet(map) %>%
  addTiles() %>%
  addPolygons(data = map, 
              fillColor = ~palette(RRB), 
              fillOpacity = 0.7, 
              color = "gray", 
              weight = 1,
              label = ~paste0("RR (B): ", RRB)) %>%
  addLegend(pal = palette, values = ~RRB, title = "RR (B)")

# Display the maps





browsable(
  tagList(
    div(
      style = "display: inline-block; width: 48%; margin-right: 3%;",
      leaflet_scr
    ),
    div(
      style = "display: inline-block; width: 48%;",
      leaflet_sbr
    )
  )
)


# cd ~/Desktop/files/hepatite/dash/dash
# git init
# git add .
# git commit -m "t dashboard"
# git push -u origin main


```



