---
title: "Hepatitis C in Saudi Arabia (2019-2022)"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The relative risk maps show a clear decrease in risk, as the maps transition from yellow shades, indicating a high relative risk, to darker purple shades, representing a low relative risk. Notably, the Makkah region, located in the western part of the
map, consistently shows the highest risk of hepatitis C across the years. In contrast, the
regions in the northern part of the map demonstrates the lowest risk during the same
time frame. 

```{r, echo=FALSE, warning=FALSE, message=FALSE, comment=NA, fig.height=4, fig.width=9.5}


setwd("~/Desktop/files/hepatite/dash/dash")


library(sf)
library(sp)
library(raster)
library(readxl)
library(mapview)
library(dplyr)
library(ggplot2)
library(patchwork)
library(viridis)
library(openair)
library(terra)
library(leaflet)
library(tigris)
library(pander)
library(spdep)
library(INLA)
library(gridExtra)
#library(rgeoboundaries)
library(rnaturalearth)
library(tidyr)
library(htmltools)



#################################
# Get the map of Saudi Arabia
#################################


map <- ne_states(country = "Saudi Arabia", returnclass = "sf")
map <- st_set_crs(map, 4326) 
map <- map[order(map$name),]
maplatlong <- st_transform(map, 4326)


map$name[1]="Asir"
map$name[2]="Al Bahah"
map$name[3]="Northern Borders"
map$name[4]="Al Jawf"
map$name[5]="Madinah"
map$name[6]="Qassim"
map$name[7]="Riyadh"
map$name[8]="Eastern Province"
map$name[9]="Hail"
map$name[10]= "Jazan"
map$name[11]="Makkah"
map$name[12]="Najran"
map$name[13]="Tabuk"






#################################
# Population of Saudi Arabia
#################################



# # # Sum population in regions of map



population <- data.frame(
  regions = c("Asir", "Al Bahah", "Northern Borders", "Al Jawf", "Madinah", "Qassim", "Riyadh", 
              "Eastern Province", "Hail", "Jazan", "Makkah", "Najran", "Tabuk"),
  pop22 = c(2024285, 339174, 373577, 595822, 2137983, 1336179, 8591748, 5125254, 746406, 1404997, 8021463, 592300, 886036),
  pop21 = c(1943532, 327833, 359411, 574894, 2053240, 1289032, 8175378, 4879962, 715422, 1355099, 7692188, 567533, 850859),
  pop20 = c(1958677, 331305, 361607, 579446, 2096074, 1310168, 8422408, 4989327, 719987, 1370141, 7982529, 574344, 856497),
  pop19 = c(1876328, 319257, 346525, 557379, 2004830, 1260269, 7979262, 4740235, 688236, 1315117, 7607556, 548237, 820568)
)



############################################################
# HepatitisC data (Areal data) 2019,2020,2021,2022 
############################################################

# read the HepatitisC data

data <- read.csv("~/Desktop/files/hepatite/dash/dash/data/HepatitisC.csv", header = T)

d <- aggregate(
  x = data$cases,
  by = list(region = data$regions, year = data$year),
  FUN = sum
)

names(d) <- c("region", "year", "Y")


data <- data[order(
  data$regions,
  data$year
), ]





population_data_long <- reshape(population, varying = list(c("pop19", "pop20", "pop21", "pop22")), 
                                v.names = "population", timevar = "year", times = c(2019, 2020, 2021, 2022), direction = "long")

population_data_long$year <- population_data_long$year

# Merge the cases and population data

data <- merge(data, population_data_long[, c("regions", "year", "population")], by = c("regions", "year"), all.x = TRUE)





###########################
# Expected value 
###########################

library(SpatialEpi)
n.strata <- 1
E <- expected(
  population = data$population,
  cases = data$cases,
  n.strata = n.strata
)

nyears <- length(unique(data$year))
regionsE <- rep(unique(data$regions),
                 each = nyears)


nregions <- length(unique(data$regions))
yearsE <- rep(unique(data$year),
              times = nregions)

dE <- data.frame(region = regionsE, year = yearsE, E = round(E))




d <- merge(d, dE, by = c("region", "year"))



d$SIR <- d$Y / d$E



dw <- reshape(d,
              timevar = "year",
              idvar = "region",
              direction = "wide"
)


map <- merge(map, dw, by.x = "name", by.y = "region")



map_sf <- gather(map, year, SIR, paste0("SIR.", 2019:2022))

map_sf$year <- as.integer(substring(map_sf$year, 5, 8))



# Create a color palette for SIR
pal_sir <- colorNumeric(palette = "viridis", domain = map_sf$SIR, na.color = "transparent")

# Initialize the leaflet map
leaflet_map <- leaflet() %>%
  addTiles() %>%
  setView(lng = 45, lat = 23.5, zoom = 4) # Center on Saudi Arabia

# Split data by year for creating individual layers
map_layers <- split(map_sf, map_sf$year)

# Add yearly SIR polygons to the map
for (year in names(map_layers)) {
  year_data <- map_layers[[year]]
  
  leaflet_map <- leaflet_map %>%
    addPolygons(
      data = year_data,
      fillColor = ~pal_sir(SIR),
      color = "gray",
      weight = 1,
      opacity = 1,
      fillOpacity = 0.7,
      popup = ~paste0(
        "<b>Region:</b> ", name, "<br>",
        "<b>Year:</b> ", year, "<br>",
        "<b>SIR:</b> ", round(SIR, 2)
      ),
      group = paste0("SIR - ", year)
    )
}

# Add layers control for toggling years, with only the first year checked by default
# leaflet_map <- leaflet_map %>%
#   addLayersControl(
#     overlayGroups = paste0("SIR - ", names(map_layers)),
#     options = layersControlOptions(collapsed = FALSE, autoZIndex = FALSE)
#   ) %>%
#   hideGroup(paste0("SIR - ", names(map_layers)[-1])) %>%
#   addLegend(
#     position = "bottomright",
#     pal = pal_sir,
#     values = map_sf$SIR,
#     title = "SIR",
#     opacity = 1
#   )

# Plot the leaflet map







library(INLA)
library(spdep)
nb <- poly2nb(map)
nb2INLA("map.adj", nb)
g <- inla.read.graph(filename = "map.adj")


d$idarea <- as.numeric(as.factor(d$region))
d$idarea1 <- d$idarea
d$idtime <- 1 + d$year - min(d$year)


formula <- Y ~ f(idarea, model = "bym", graph = g) +
  f(idarea1, idtime, model = "iid") + idtime


res <- inla(formula,
            family = "poisson", data = d, E = E,
            control.predictor = list(compute = TRUE),
            control.compute = list(dic = T, waic = T, cpo = T)
)




d$RR <- res$summary.fitted.values[, "mean"]
d$LL <- res$summary.fitted.values[, "0.025quant"]
d$UL <- res$summary.fitted.values[, "0.975quant"]

map_sf <- merge(
  map_sf, d,
  by.x = c("name", "year"),
  by.y = c("region", "year")
)



# Create a color palette for RR
pal_rr <- colorNumeric(palette = "viridis", domain = map_sf$RR, na.color = "transparent")

# Initialize the leaflet map for RR
leaflet_map_rr <- leaflet() %>%
  addTiles() %>%
  setView(lng = 45, lat = 23.5, zoom = 4) # Center on Saudi Arabia

# Split data by year for creating individual layers
map_layers_rr <- split(map_sf, map_sf$year)

# Add yearly RR polygons to the map
for (year in names(map_layers_rr)) {
  year_data <- map_layers_rr[[year]]
  
  leaflet_map_rr <- leaflet_map_rr %>%
    addPolygons(
      data = year_data,
      fillColor = ~pal_rr(RR),
      color = "gray",
      weight = 1,
      opacity = 1,
      fillOpacity = 0.7,
      popup = ~paste0(
        "<b>Region:</b> ", name, "<br>",
        "<b>Year:</b> ", year, "<br>",
        "<b>RR:</b> ", round(RR, 2)
      ),
      group = paste0("RR - ", year)
    )
}

# Add layers control for toggling years, with only the first year checked by default
leaflet_map_rr <- leaflet_map_rr %>%
  addLayersControl(
    overlayGroups = paste0("RR - ", names(map_layers_rr)),
    options = layersControlOptions(collapsed = FALSE, autoZIndex = FALSE)
  ) %>%
  hideGroup(paste0("RR - ", names(map_layers_rr)[-1])) %>%
  addLegend(
    position = "bottomright",
    pal = pal_rr,
    values = map_sf$RR,
    title = "RR",
    opacity = 1
  )

leaflet_map_rr
# Plot the leaflet map for RR





# browsable(
#   tagList(
#     div(
#       style = "display: inline-block; width: 48%; margin-right: 3%;", 
#       leaflet_map
#     ),
#     div(
#       style = "display: inline-block; width: 48%;", 
#       leaflet_map_rr
#     )
#   )
# )



```

